\section*{The People behind the Book}

\subsubsection*{Thomas Hunter II, Author}

\href{http://thomashunter.name}{Thomas} previously worked as a Developer Advocate and API Architect of a large file-sharing and storage service, where his main concern was getting a well-documented API into the hands of third-party developers. While working at a company in the financial industry he evangelized an internal API façade to abstract backend disparity. He's also given various talks on the topic of API Design. His previous book \href{http://amzn.to/1aqVRvq}{Backbone.js Application Development} guides the reader through the process of building an API-consuming JavaScript application.

\subsubsection*{John Sheehan, Technical Reviewer}

John is an API fanatic with over 15 years of experience working in a wide variety of IT and software development roles. As an early employee at Twilio, John lead the developer evangelism program and worked as a Product Manager for Developer Experience. After Twilio, John was Platform Lead at IFTTT working with API providers to create new channels. John is also the creator of \href{https://www.github.com/restsharp/restsharp}{RestSharp}, \href{http://www.apidigest.com/}{API Digest}, \href{http://www.api-jobs.com/}{API Jobs} and co-host of \href{http://trafficandweather.io/}{Traffic and Weather}, an API and cloud podcast.

\subsubsection*{Jon Kuperman, Influence}

I'd like to make a special thanks to \href{https://twitter.com/thecodeplanet}{Jon}, who once told me "Tom, you should turn that \href{http://codeplanet.io/principles-good-restful-api-design/}{API blog post} into a book."

\newpage
